---
title: "PP_project"
author: "Loman Sezestre, Florian Jacquetin"
format:
  html:
    code-fold: false
execute:
  echo: true
  output: true
  warning: false
  message: false
---

#| setup
#| include: true
#| echo: false

# Solving the price puzzle

This project aims at studying the price puzzle, a macroeconomic paradox observed by Sims (1980), when an increase of interest rate is sometimes followed by an increase of inflation in macroeconometric models.

It is designed as follows :

1.  Presentation of the data
2.  Reminder of the results of the NK framework
3.  Modeling of the MP in US according to Castelnuovo and Surico (2010)
4.  Solving the price puzzle according different strategies
5.  Extensions

```{r}

packages <- c(
  "tidyverse",
  "lubridate",
  "zoo",
  "fredr",
  "openxlsx",
  "mFilter",
  "R.matlab",
  "strucchange",
  "vars"
)

to_install <- packages[!packages %in% installed.packages()[, "Package"]]
if(length(to_install) > 0) install.packages(to_install)

for(p in packages) library(p, character.only = TRUE)

source("utils.R")
```

Output gap (<https://www.cbo.gov/data/budget-economic-data>)

```{r}
ogap <- read.csv2("data/Quarterly_January2025.csv") %>% 
  mutate(
    date = as.yearqtr(date, format = "%Y q%q")
  )
```

```{r}
#| echo: false
ogap <- read.csv2("data/Quarterly_January2025.csv") %>% 
  mutate(
    date = gsub("q","Q",date),
    date = as.yearqtr(date, format = "%Y Q%q"),
    y_gap = as.numeric(output_gap)
  ) %>%
  dplyr::select(date,y_gap)

ggplot(ogap, aes(x = as.numeric(date), y = y_gap)) +
  theme_minimal() +
  geom_line() +
  labs(x = NULL, y = "Output Gap")
```

GDP Deflator and Inflation (FREDR)

```{r}
fredr::fredr_set_key("2081d90b146ea654a4866e909178bfaf")

gdpdef <- fredr::fredr(
  series_id = "GDPDEF",
  observation_start = as.Date("1960-01-01")
) %>%
  transmute(
    date = as.yearqtr(date),
    p_gdp = log(value)
  ) %>%
  arrange(date)

pi_gdp <- gdpdef %>%
  mutate(
    pi = 400 * (p_gdp - lag(p_gdp))    # inflation trimestrielle annualisée
  ) %>%
  drop_na() %>%
  dplyr::select(date, pi)

ggplot(pi_gdp, aes(date, pi)) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL, y = "Quarterly inflation (annualized)")
```

```{r}
ffr_m <- fredr::fredr(
  series_id = "FEDFUNDS",
  observation_start = as.Date("1960-01-01")
) %>%
  transmute(
    date_m = as.yearmon(date),
    ffr = value
  )

ffr_q <- ffr_m %>%
  mutate(
    date = as.yearqtr(date_m)
  ) %>%
  group_by(date) %>%
  summarise(
    ffr = mean(ffr, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(date) 
# %>%
# filter(date <= "2006 Q4")

ggplot(ffr_q, aes(date, ffr)) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL, y = "Federal Funds Rate")
```
We do a Chow test to highlight the change of regime post 1979.


```{r}
df_us <- ogap %>%
  inner_join(pi_gdp, by = "date") %>%
  inner_join(ffr_q,  by = "date") %>%
  arrange(date)


print(head(df_us, 10))

# 6. Sauvegarde --

dir.create("sorties", showWarnings = FALSE)
```
We run a Chow test to find the best break point, following Paul Volcker's appointment at the Fed.

```{r}
df <- df_us %>% dplyr::select(-date)

select_order <- VARselect(df)
lags <- select_order$selection["SC(n)"]

var_fit <- VAR(df, p = lags, type = "const")
ffr_eq <- var_fit$varresult$ffr
resid_ffr <- residuals(ffr_eq)

quarters <- seq(as.yearqtr("1979 Q1"), as.yearqtr("1981 Q4"), by = 0.25)

p_values <- numeric(length(quarters))

for(i in seq_along(quarters)) {
  bp <- which(df_us$date == quarters[i])
  if(length(bp) == 1) { 
    chow <- sctest(resid_ffr ~ 1, type = "Chow", point = bp)
    p_values[i] <- chow$p.value
  } else {
    p_values[i] <- NA
  }
}

p_df <- data.frame(
  Quarter = quarters,
  P_value = p_values
)

ggplot(p_df, aes(x = Quarter, y = P_value)) +
  geom_line(color = "red", size = 1.2) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "blue") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Chow test p-values",
    x = NULL,
    y = "p-value"
  )
```
```{r}
select_order <- VARselect(df)
lags <- select_order$selection["SC(n)"]

var_fit <- VAR(df, p = lags, type = "const")
ffr_eq <- var_fit$varresult$ffr
resid_ffr <- residuals(ffr_eq)

quarters <- seq(as.yearqtr("2004 Q1"), as.yearqtr("2011 Q4"), by = 0.25)

p_values <- numeric(length(quarters))

for(i in seq_along(quarters)) {
  bp <- which(df_us$date == quarters[i])
  if(length(bp) == 1) { 
    chow <- sctest(resid_ffr ~ 1, type = "Chow", point = bp)
    p_values[i] <- chow$p.value
  } else {
    p_values[i] <- NA
  }
}

p_df <- data.frame(
  Quarter = quarters,
  P_value = p_values
)

ggplot(p_df, aes(x = Quarter, y = P_value)) +
  geom_line(color = "red", size = 1.2) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "blue") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Chow test p-values",
    x = NULL,
    y = "p-value"
  )
```


Here we introduce the NK model and its properties.

```{r} 
#| echo: false
## This cell needs to have Matlab installed. Change the link if necessary

matlab_exec <- "C:/Program Files/MATLAB/R2024b/bin/matlab.exe"

system(sprintf('"%s" -batch "run(\'compute_nk_model.m\')"', matlab_exec))
```

The parameters are taken from Clarida, R., Galí, J., & Gertler, M. (2000).

Let's observe the IRFs.

```{r}

mat_data <- R.matlab::readMat("sorties/irfs.mat")
oo_matrix <- mat_data$oo.       
irfs_list <- oo_matrix[[20]]

x_eb <- as.numeric(irfs_list[[6]])
pi_eb <- as.numeric(irfs_list[[7]])
int_eb <- as.numeric(irfs_list[[8]])

nk_df <- data.frame(
  horizon = 0:(length(x_eb)-1),
  y_gap = x_eb,
  pi = pi_eb,
  ffr = int_eb,
  check.names = FALSE
) %>%
  pivot_longer(cols = -horizon, names_to = "variable", values_to = "irf")

ggplot(nk_df, aes(x = horizon, y = irf)) +
  theme_minimal(base_size = 14) +
  geom_line(size = 1, color = "red") +
  facet_wrap(~variable, ncol = 1, scales = "free_y") +
  labs(title = "IRFs to a MP shock in a NK model",
       x = "Periods",
       y = "Response") +
  geom_hline(yintercept = 0, color = "grey", linetype = "solid") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "solid") +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 12),
    panel.grid.major = element_line(color = "gray90")
  )


```

Let's construct a SVAR model based on [og(t) pi(t) ffr(t)]

```{r}
# echo: false
Y_pre_ts  <- df_us %>%
  filter(date <= as.yearqtr("1979 Q4")) %>%
  df_to_ts()

Y_post_ts <- df_us %>%
  filter(date >= as.yearqtr("1980 Q1"),
         date <= as.yearqtr("2006 Q2")) %>%
  df_to_ts()

Y_end_ts <- df_us %>%
  filter(date >= as.yearqtr("2009 Q2"),
         date <= as.yearqtr("2019 Q4")) %>%
  df_to_ts()

lags_pre  <- VARselect(Y_pre_ts,  lag.max = 8, type = "const")
lags_post <- VARselect(Y_post_ts, lag.max = 8, type = "const")
lags_end <- VARselect(Y_end_ts, lag.max = 8, type = "const")

p_pre  <- lags_pre$selection["SC(n)"]
p_post <- lags_post$selection["SC(n)"]
p_end <- lags_end$selection["SC(n)"]

cat("Lags retenus (pré)  :", p_pre,  "\n")
cat("Lags retenus (post) :", p_post, "\n")
cat("Lags retenus (post) :", p_end, "\n")

var_pre <- VAR(Y_pre_ts,  p = p_pre,  type = "const")
var_post <- VAR(Y_post_ts, p = p_post, type = "const")
var_end <- VAR(Y_end_ts, p = p_post, type = "const")

summary(var_pre)
summary(var_post)

# 7. IRFs with Cholesky identification 

irf_pre <- irf(
  var_pre,
  impulse = "ffr",
  response = c("y_gap", "pi", "ffr"),
  n.ahead = 20,
  boot = TRUE,
  runs = 500,
  ortho = TRUE  # Cholesky
)


irf_post <- irf(
  var_post,
  impulse = "ffr",
  response = c("y_gap", "pi", "ffr"),
  n.ahead = 20,
  boot = TRUE,
  runs = 500,
  ortho = TRUE
)

irf_end <- irf(
  var_end,
  impulse = "ffr",
  response = c("y_gap", "pi", "ffr"),
  n.ahead = 20,
  boot = TRUE,
  runs = 500,
  ortho = TRUE
)
df_irf_pre  <- irf_to_df(irf_pre, impulse_name = "ffr")
df_irf_post <- irf_to_df(irf_post, impulse_name = "ffr")
df_irf_end <- irf_to_df(irf_end, impulse_name = "ffr")

df_irf_pre$period  <- factor("Pré-1979", levels = c("Pré-1979", "Post-1979","Post-2007"))
df_irf_post$period <- factor("Post-1979", levels = c("Pré-1979", "Post-1979","Post-2007"))
df_irf_end$period <- factor("Post-2007", levels = c("Pré-1979", "Post-1979","Post-2007"))

nk_df_long <- nk_df %>%
  mutate(lower = irf, upper = irf) %>%   # bandes de confiance égales à la valeur
  expand_grid(period = c("Pré-1979", "Post-1979","Post-2007"))

df_irf_all <- bind_rows(df_irf_pre, df_irf_post,df_irf_end)


# Ajouter un identifiant pour la légende combinée
df_irf_all_plot <- df_irf_all %>%
  mutate(period_factor = factor(period, levels = c("Pré-1979", "Post-1979","Post-2007")))

nk_df_long_plot <- nk_df_long %>%
  mutate(period_factor = factor(period, levels = c("Pré-1979", "Post-1979","Post-2007")))

ggplot() +
  geom_ribbon(data = df_irf_all_plot,
              aes(x = horizon, ymin = lower, ymax = upper, fill = period_factor),
              alpha = 0.15, color = NA, show.legend = FALSE) +
  geom_line(data = df_irf_all_plot,
            aes(x = horizon, y = irf, color = period_factor),
            size = 1, show.legend = FALSE) +
  geom_line(data = nk_df_long_plot,
            aes(x = horizon, y = irf, linetype = "NK model"),
            color = "black", size = 1) +
  facet_grid(variable ~ period_factor, scales = "free_y") +
  scale_color_manual(values = c("Pré-1979" = "red", "Post-1979" = "blue","Post-2007" = "green")) +
  scale_fill_manual(values = c("Pré-1979" = "red", "Post-1979" = "blue", "Post-2007" = "green")) +
  scale_linetype_manual(name = "", values = c("NK model" = "dotted")) +
  labs(title = "IRFs to ffr shock", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
  plot.title = element_text(face = "bold", hjust = 0.5,size = 12),
    strip.text = element_text(face = "bold")
  )

```

We want to illustrate where the price puzzle comes from before 1979 and after 2007. We plot the Forecast Error Variance Decomposition (FEVD) after the ffr shock.


```{r}
# FEVD pour le VAR à 3 variables ----------------------------------------

fevd_pre  <- fevd(var_pre,  n.ahead = 20)
fevd_post <- fevd(var_post, n.ahead = 20)
fevd_end  <- fevd(var_end,  n.ahead = 20)

fevd_to_df <- function(fevd_obj, var_name, period_label) {
  mat <- fevd_obj[[var_name]]           # matrice horizons x variables
  df  <- as.data.frame(mat)
  df$horizon <- 0:(nrow(df) - 1)
  
  df %>%
    pivot_longer(
      cols = -horizon,
      names_to = "shock",
      values_to = "share"
    ) %>%
    mutate(period = period_label)
}

fevd_pi_all <- bind_rows(
  fevd_to_df(fevd_pre,  "pi", "Pré-1979"),
  fevd_to_df(fevd_post, "pi", "Post-1979"),
  fevd_to_df(fevd_end,  "pi", "Post-2007")
) %>%
  mutate(
    period = factor(period, levels = c("Pré-1979", "Post-1979", "Post-2007"))
  )

ggplot(fevd_pi_all, aes(x = horizon, y = share, fill = shock)) +
  geom_area(position = "stack") +
  facet_wrap(~ period, nrow = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Forecast error variance decomposition of inflation",
    x = "Quarters",
    y = "Share of forecast variance (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```
Finally, let's see the estimated coefficients of the VAR.

```{r}
eq_pre_pi  <- summary(var_pre)$varresult$pi

coef_pre  <- coef(eq_pre_pi)

print(coef_pre)

```
Before 1979, the ffr coefficient is positive and large. Before Volcker, the Fed typically raised interest rates in response to expected increases in inflation (poor monetary policy rules, focus on output and unemployment, weak anti-inflation credibility). The VAR confuses cause and consequence.

No Phillips curve effect : output gap has no effect on inflation.

```{r}
eq_post_pi  <- summary(var_post)$varresult$pi

coef_post  <- coef(eq_post_pi)

print(coef_post)
```
Here no short effect, only a medium negative and significant + functioning Phillips curve.

After 1979, the Fed aggressively fought inflation (Volcker disinflation),

```{r}
eq_end_pi  <- summary(var_end)$varresult$pi

coef_end  <- coef(eq_end_pi)

print(coef_end)
```
breakdown of the standard VAR structure. Coefficients on interest rates are huge and unstable. It is typical of the ZLB / QE environment.

Apparent but spurious price puzzle.

# Trying the fix the puzzle

Following Castelnuovo et al., we propose a sign restrictions approach. We impose that after a monetary shock, ffr > O and outputggap < 0. Inflations remains unconstrained.

```{r}
Sigma <- cov(resid(var_pre))
P <- t(chol(Sigma))
n.sim <- 5000
H <- 12
N <- ncol(Y_pre_ts)

A_raw <- vars::Bcoef(var_pre)

if (is.list(A_raw)) {
  A <- do.call(cbind, A_raw)
} else {
  A <- A_raw
}

p <- var_pre$p
A_big <- matrix(0, nrow = N * p, ncol = N * p)

A_no_const <- A[, 1:(N * p)]
A_big[1:N, 1:(N * p)] <- A_no_const

if (p > 1) {
  A_big[(N+1):(N*p), 1:(N*(p-1))] <- diag(N * (p-1))
}

irfs_kept <- list()
count <- 0
set.seed(123)

for (i in 1:n.sim) {
  Z <- matrix(rnorm(N*N), N, N)
  Q <- qr.Q(qr(Z))
  B <- P %*% Q
  irf_B <- manual_irf(B, H)
  dy_response   <- irf_B[1, "y_gap"]
  taux_response <- irf_B[1, "ffr"]
  
  if (taux_response >= 0 && dy_response <= 0) {
    count <- count + 1
    irfs_kept[[count]] <- irf_B
  }
}

percent_kept <- count / n.sim * 100
cat(percent_kept, "% of accepted simulations\n")

### --- 4. Tidy-data des IRFs retenues ----

irf_df <- purrr::map_dfr(
  irfs_kept,
  ~as.data.frame(.x) %>% mutate(h = 1:H),
  .id = "sim"
) %>%
  tidyr::pivot_longer(
    cols = -c(sim, h),
    names_to = "variable",
    values_to = "irf"
  )

### --- 5. Statistiques (mean & quantiles) ----

irf_summary <- irf_df %>%
  group_by(h, variable) %>%
  summarise(
    mean    = mean(irf),
    lower68 = quantile(irf, 0.16),
    upper68 = quantile(irf, 0.84),
    lower90 = quantile(irf, 0.05),
    upper90 = quantile(irf, 0.95),
    .groups = "drop"
  )

### --- 6. IRFs Cholesky pré-1979 (benchmark VAR) ----

# Extraction correcte : matrice H x 3
chol_mat <- irf_pre_chol_raw$ffr    # matrice 12 x 3
var_names <- colnames(chol_mat)     # "y_gap" "pi" "ffr"

# Construction du tidy dataframe
irf_pre_df <- bind_rows(
  lapply(1:N, function(j){
    tibble(
      variable = var_names[j],
      h = 1:H,
      irf = chol_mat[1:H, j]
    )
  })
)

### --- 7. GRAPHIQUE FINAL ----

ggplot() +
  geom_ribbon(data=irf_summary,aes(x=h,ymin=lower90,ymax=upper90),alpha=.15,fill="lightblue") +
  geom_ribbon(data=irf_summary,aes(x=h,ymin=lower68,ymax=upper68),alpha=.3,fill="blue") +
  geom_line(data=irf_summary,aes(x=h,y=mean,color="Sign restrictions",linetype="Sign restrictions"),size=1.2) +
  geom_line(data=irf_pre_df,aes(x=h,y=irf,color="Baseline",linetype="Baseline"),size=1) +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  facet_wrap(~variable,scales="free_y") +
  labs(x="Quarters",y="IRF",title="Sign restrictions",color=NULL,linetype=NULL) +
  scale_color_manual(values=c("Sign restrictions"="black","Baseline"="red")) +
  scale_linetype_manual(values=c("Sign restrictions"="solid","Baseline"="dashed")) +
  theme_minimal(base_size=13) +
  theme(legend.position="bottom")




```

In VARs, supply and demand shocks are mixed. Before 1979, many shocks are supply shocks : both inflation and ffr increase, and output gap decreases. Because response of inflation to output gap is positive, removing these positive responses reduce the level of inflation. But it doesn't remove the price puzzle.

We propose now to augment the VAR. As suggested by Casteluovo, we get the inflation forecasts from the Greenbook.

```{r}
# ==============================
# 1. Charger données Greenbook
# ==============================
green <- read.xlsx("data/greenbook.xlsx")

df <- df_pre %>%
  mutate(date = gsub(" ", "", date)) %>%
  inner_join(green, by = "date")

df_inf <- df %>%
  dplyr::select(date, pi, green) %>%
  mutate(
    pi = as.numeric(pi),
    green = as.numeric(green),
    date = as.yearqtr(date, format = "%YQ%q")
  )

ggplot(df_inf, aes(x = date)) +
  theme_minimal() +
  geom_line(aes(y = pi, color = "pi"), linewidth = 1) +
  geom_line(aes(y = green, color = "exp"), linewidth = 1) +
  scale_color_manual(values = c("pi" = "blue", "exp" = "darkgreen")) +
  labs(
    x = "Date",
    y = "Value",
    color = "Variable",
    title = "Inflation vs Expectations (Greenbook)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# ==============================
# 2. VAR avec anticipations
# ==============================
df_ts <- df_to_ts(df)
lags_exp <- VARselect(df_ts, lag.max = 8, type = "const")$selection["SC(n)"]
VAR_exp <- VAR(df_ts, p = lags_exp, type = "const")


# ==============================
# 3. IRFs anticipations
# ==============================
irf_exp <- irf(
  VAR_exp,
  impulse = "ffr",
  response = c("green", "y_gap", "pi", "ffr"),
  n.ahead = 20,
  boot = TRUE,
  runs = 500,
  ortho = TRUE
)

irf_exp_df <- irf_to_df(irf_exp, impulse_name = "ffr") %>%
  mutate(model = "Avec anticipations")


# ==============================
# 4. IRF baseline (déjà calculé)
# ==============================
irf_pre_df <- irf_to_df(irf_pre, impulse_name = "ffr") %>%
  mutate(model = "Baseline")

# Ajouter un panel 'green' vide pour aligner le facet_wrap
green_missing <- irf_pre_df %>%
  filter(variable == "pi") %>%     # duplique la structure
  mutate(
    variable = "green",
    irf = NA, lower = NA, upper = NA
  )

irf_pre_df <- bind_rows(irf_pre_df, green_missing)


# ==============================
# 5. Fusion tidy
# ==============================
irf_all <- bind_rows(irf_exp_df, irf_pre_df) %>%
  mutate(model = factor(model, levels = c("Baseline", "Avec anticipations")))


# ==============================
# 6. Figure finale
# ==============================
ggplot(irf_all, aes(x = horizon, y = irf, color = model, linetype = model)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = model),
              alpha = 0.15, color = NA) +
  geom_line(size = 1.1) +
  facet_wrap(~variable, scales = "free_y") +
  scale_color_manual(values = c("Baseline" = "red",
                                "Avec anticipations" = "black")) +
  scale_fill_manual(values = c("Baseline" = "red",
                               "Avec anticipations" = "black")) +
  scale_linetype_manual(values = c("Baseline" = "dashed",
                                   "Avec anticipations" = "solid")) +
  labs(
    x = "Horizon",
    y = "IRF",
    title = "IRFs to Monetary Policy Shock: Baseline vs Inflation Expectations (Greenbook)",
    color = NULL,
    linetype = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

```


VAR à 4 variables 
```{r}

# Charger la série mensuelle WTI 
oil_m <- fredr::fredr(
  series_id = "PPIIDC", #PPIACO
  observation_start = as.Date("1960-01-01")
) %>%
  transmute(
    date_m = as.yearmon(date),
    oil    = value
  ) %>%
  drop_na(oil)

oil_q <- oil_m %>%
  mutate(date = as.yearqtr(date_m)) %>%
  group_by(date) %>%
  summarise(oil = mean(oil), .groups = "drop") %>%
  mutate(
    oil_log = log(oil),
    poil    = 400 * (oil_log - dplyr::lag(oil_log))   # inflation du pétrole
  ) %>%
  dplyr::select(date, poil) %>%
  drop_na()

# Fusion avec données US (y_gap, pi, ffr)
df_us_oil <- df_us %>%
  inner_join(oil_q, by = "date") %>%
  arrange(date) %>%
  dplyr::select(date,poil,y_gap,pi,ffr)

# Définir les 3 sous-périodes
df_pre_oil <- df_us_oil %>% filter(date <= as.yearqtr("1979 Q4"))
df_post_oil <- df_us_oil %>% filter(date >= as.yearqtr("1980 Q1"),
                                    date <= as.yearqtr("2006 Q2"))
df_end_oil <- df_us_oil %>% filter(date >= as.yearqtr("2009 Q4"),
                                   date <= as.yearqtr("2019 Q4"))

Y_pre_ts_oil  <- df_to_ts(df_pre_oil)
Y_post_ts_oil <- df_to_ts(df_post_oil)
Y_end_ts_oil  <- df_to_ts(df_end_oil)

# Choix des lags
p_pre_oil  <- VARselect(Y_pre_ts_oil,  lag.max = 8)$selection["SC(n)"]
p_post_oil <- VARselect(Y_post_ts_oil, lag.max = 8)$selection["SC(n)"]
p_end_oil  <- VARselect(Y_end_ts_oil,  lag.max = 16)$selection["SC(n)"]

# Estimation VAR
var_pre_oil  <- VAR(Y_pre_ts_oil,  p = p_pre_oil,  type = "const")
var_post_oil <- VAR(Y_post_ts_oil, p = p_post_oil, type = "const")
var_end_oil  <- VAR(Y_end_ts_oil,  p = p_end_oil,  type = "const")

# IRFs Cholesky
irf_pre_oil  <- irf(var_pre_oil,  impulse="ffr", response=c("y_gap","pi","poil","ffr"),
                    n.ahead=20, boot=TRUE, runs=500, ortho=TRUE)
irf_post_oil <- irf(var_post_oil, impulse="ffr", response=c("y_gap","pi","poil","ffr"),
                    n.ahead=20, boot=TRUE, runs=500, ortho=TRUE)
irf_end_oil  <- irf(var_end_oil,  impulse="ffr", response=c("y_gap","pi","poil","ffr"),
                    n.ahead=20, boot=TRUE, runs=500, ortho=TRUE)

# Mise en forme tidy
df1 <- irf_to_df(irf_pre_oil, "ffr")  %>% mutate(period = "Pré-1979")
df2 <- irf_to_df(irf_post_oil, "ffr") %>% mutate(period = "Post-1979")
df3 <- irf_to_df(irf_end_oil, "ffr")  %>% mutate(period = "Post-2007")

df_irf_oil <- bind_rows(df1, df2, df3) %>%
  mutate(period = factor(period, levels=c("Pré-1979","Post-1979","Post-2007")))

# Affichage clean
p_oil <- ggplot(df_irf_oil) +
  geom_ribbon(aes(x = horizon, ymin = lower, ymax = upper, fill = period),
              alpha = 0.15, color = NA) +
  geom_line(aes(x = horizon, y = irf, color = period), size = 1) +
  facet_grid(variable ~ period, scales = "free_y") +
  scale_color_manual(values = c("Pré-1979" = "red",
                                "Post-1979" = "blue",
                                "Post-2007" = "green")) +
  scale_fill_manual(values = c("Pré-1979" = "red",
                               "Post-1979" = "blue",
                               "Post-2007" = "green")) +
  labs(title = "IRFs to a monetary policy shock (VAR with Oil)",
       x = "Horizon (quarters)", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text  = element_text(face = "bold"),
    plot.title  = element_text(face = "bold", hjust = 0.5)
  )

print(p_oil)
```
